---
title: "data vis final"
output: html_document
---

In our project, we wanted to see how density and overcrowding in NYC affected covid rates. Below we did a tidycensus API call to collect zip code-level data from the ACS 2015-2019. Then we compared that with covid data from the city, which can be accessed here: https://data.cityofnewyork.us/dataset/DOHMH-COVID-19-Antibody-by-Modified-ZIP-Code-Tabul/6qs8-44ki

```{r}
library(tidycensus)
library(tidyverse)
library(janitor)
library(ggpubr)
```


```{r}
acs_vars <- load_variables(2019, "acs5", cache = TRUE)
```

Even though we only wanted to cover crowding and density, we also wanted to cover our bases and do controls on those variables, so we included a bunch of different variables in the call.

```{r}
zip_acs_raw <- get_acs(
  geography = "zcta",
  state = "NY",
  variables = c(
    "population" = "B01001_001",
    "white" = "B02001_002",
    "black" = "B02001_003",
    "asian" = "B02001_005",
    "pacific_islander" = "B02001_006",
    "hispanic" = "B03001_003",
    "total_units" = "B25001_001",
    "units_occupied" = "B25002_002",
    "units_vacant" = "B25002_003",
    "owner_occupied" = "B25003_002",
    "renter_occupied" = "B25003_003",
    "crowding_total" = "B25014_001",
    "owner_crowding_total" = "B25014_002",
    "owner_1_1.5_per_room" = "B25014_005",
    "owner_1.5_2_per_room" = "B25014_006",
    "owner_over_2_per_room" = "B25014_007",
    "renter_crowding_total" = "B25014_008",
    "renter_1_1.5_per_room" = "B25014_011",
    "renter_1.5_2_per_room" = "B25014_012",
    "renter_over_2_per_room" = "B25014_013",
    "units_in_structure_total" = "B25024_001",
    "detached" = "B25024_002",
    "attached" = "B25024_003",
    "10_19_unit_structure" = "B25024_007",
    "20_49_unit_structure" = "B25024_008",
    "over_50_unit_structure" = "B25024_009",
    "household_vehicles_total" = "B25044_001",
    "owner_occupied_no_cars" = "B25044_003",
    "renter_occupied_no_cars" = "B25044_010",
    "median_gross_rent" = "B25064_001",
    "rent_percent_hh_income" = "B25071_001",
    "rent_burden_total" = "B25070_001",
    "rent_burden_30_35_percent" = "B25070_007",
    "rent_burden_35_40_percent" = "B25070_008",
    "rent_burden_40_50_percent" = "B25070_009",
    "rent_burden_over_50_percent" = "B25070_010",
    "median_household_income" = "B19013_001",
    "education_total" = "B15003_001",
    "hs_diploma" = "B15003_017",
    "ged" = "B15003_018",
    "college>1" = "B15003_019",
    "college<1_no_degree" = "B15003_020",
    "associates_degree" = "B15003_021",
    "bachelors_degree" = "B15003_022",
    "masters_degree" = "B15003_023",
    "professional_degree" = "B15003_024",
    "doctorate" = "B15003_025",
    "citizenship_total" = "B05001_001",
    "naturalized_citizen" = "B05001_005",
    "non_citizen" = "B05001_006",
    "commute_total" = "B08301_001",
    "commute_by_car" = "B08301_002",
    "commute_by_public_transit" = "B08301_010",
    "commute_by_bus" = "B08301_011",
    "commute_by_subway" = "B08301_012",
    "commute_by_commuter_rail" = "B08301_013",
    "commute_by_other_rail" = "B08301_014",
    "commute_by_walk" = "B08301_019",
    "commute_by_bicycle" = "B08301_018",
    "commute_by_ferry" = "B08301_015",
    "poverty_total" = "C17002_001",
    "poverty_<.5" = "C17002_002",
    "poverty_.5_1.0" = "C17002_003"
    
  ),
  survey = "acs5",
  year = 2019,
  output = "wide",
  geometry = FALSE
)
```


```{r}
write.csv(zip_acs_raw, "zip_acs_raw.csv", row.names = FALSE)
```

I totally forgot to get vacancy data, so I snagged that later.

```{r}
zip_acs2_raw <- get_acs(
  geography = "zcta",
  state = "NY",
  variables = c(
    "vacancy_total" = "B25004_001",
    "for_rent" = "B25004_002",
    "rented_not_occ" = "B25004_003",
    "for_sale" = "B25004_004",
    "sold_not_occ" = "B25004_005",
    "seasonal_rec_occaisional" = "B25004_006",
    "migrant_worker" = "B25004_007",
    "other_vacant" = "B25004_008"
    
  ),
  survey = "acs5",
  year = 2019,
  output = "wide",
  geometry = FALSE
)
```


```{r}
write.csv(zip_acs2_raw, "zip_acs2_raw.csv", row.names = FALSE)
```

```{r}
covid_census <- read_csv("combined acs covid zip data - transformations 2.csv", 
    col_types = cols(zip = col_character(), 
        lat = col_number(), lon = col_number(), 
        covid_case_rate = col_number(), population_covid = col_number(), 
        covid_death_rate = col_number(), 
        covid_positive_percent = col_number(), 
        population_acs = col_number(), median_household_income = col_number(), 
        median_gross_rent = col_number(), 
        white = col_number(), black = col_number(), 
        aapi = col_number(), hisp = col_number(), 
        nonwhite = col_number(), renter_not_owner = col_number(), 
        vacancy_rate_for_rent_or_sale = col_number(), 
        crowding = col_number(), severe_crowding = col_number(), 
        sfh = col_number(), `10+unit_building` = col_number(), 
        `20+unit_building` = col_number(), 
        `50+unit_building` = col_number(), 
        hhs_no_cars = col_number(), rent_percent_hh_income = col_number(), 
        rent_burdened = col_number(), severely_rent_burdened = col_number(), 
        `hs_diploma+` = col_number(), `bachelors+` = col_number(), 
        foreign_born = col_number(), car_commute = col_number(), 
        public_transit = col_number(), `walk+bike` = col_number(), 
        poverty = col_number()))
```

We did the join in google sheets, so the full combined data spreadsheet is the one above. Below is a series of correlations and regressions and charts and graphs we made to try to discover the most important variables when it comes to understanding covid spread.

```{r}
plot(data = covid_census, covid_positive_percent ~ vacancy_rate_for_rent_or_sale)
```

```{r}
covid_t2 <- read_csv("combined acs covid zip data - t2.csv")
```

```{r}
covidt2 <- clean_names(covid_t2)
```


```{r}
hist(covid_census$crowding)
```

```{r}
hist(covidt2$covid_case_rate)
```

```{r}
cor(covid_census$covid_positive_percent, covid_census$aapi)
```

```{r}
cor(covidt2$covid_case_rate, covidt2$x10_b100)
```

```{r}
positive.crowding.lm <- lm(data = covid_census, covid_positive_percent ~ crowding)

summary(positive.crowding.lm)
```


```{r}
covid.car.rent.lm <- lm(covid_case_rate ~ carcommute100 + median_gross_rent, data = covid_t2)

summary(covid.car.rent.lm)
```

```{r}
covid.density.crowding.lm <- lm(covid_case_rate ~ x10_b100 + crowding100, data = covidt2)

summary(covid.density.crowding.lm)
```

```{r}
covid.density.lm <- lm(covid_case_rate ~ x10_b100, data = covidt2)

summary(covid.density.lm)
```

```{r}
covid.severecrowding.lm <- lm(covid_case_rate ~ severecrowding100, data = covidt2)

summary(covid.severecrowding.lm)
```

```{r}
covid.car.no_car.lm <- lm(covid_case_rate ~ carcommute100 + nocars100, data = covidt2)

summary(covid.car.no_car.lm)
```

```{r}
covid.foreignborn.lm <- lm(covid_case_rate ~ foreign100, data = covidt2)

summary(covid.foreignborn.lm)
```

```{r}
covid.aapi.lm <- lm(covid_case_rate ~ aapi100, data = covidt2)

summary(covid.aapi.lm)
```

```{r}
covid.crowding.lm <- lm(covid_case_rate ~ crowding100, data = covidt2)

summary(covid.crowding.lm)
```

```{r}
covid.poverty.lm <- lm(covid_case_rate ~ poverty100, data = covidt2)

summary(covid.poverty.lm)
```

```{r}
covid.vacancy.lm <- lm(covid_case_rate ~ vacancy100, data = covidt2)

summary(covid.vacancy.lm)
```

```{r}
covid.hisp.lm <- lm(covid_case_rate ~ hisp100, data = covidt2)

summary(covid.hisp.lm)
```

```{r}
covid.ba.income.lm <- lm(covid_case_rate ~ ba_100 + median_household_income, data = covidt2)

summary(covid.ba.income.lm)
```

```{r}
covid.ba.lm <- lm(covid_case_rate ~ ba_100, data = covidt2)

summary(covid.ba.lm)
```

```{r}
covid.b10.income.lm <- lm(covid_case_rate ~ x10_b100 + median_household_income, data = covidt2)

summary(covid.b10.income.lm)
```

```{r}
covid.carcommute.income.lm <- lm(covid_case_rate ~ carcommute100 + median_household_income, data = covidt2)

summary(covid.carcommute.income.lm)
```

```{r}
covid.rent.income.lm <- lm(covid_case_rate ~ median_gross_rent + median_household_income, data = covidt2)

summary(covid.rent.income.lm)
```

```{r}
covid.sfh.income.lm <- lm(covid_case_rate ~ sfh100 + median_household_income, data = covidt2)

summary(covid.sfh.income.lm)
```

```{r}
covid.carcommute.ba.lm <- lm(covid_case_rate ~ carcommute100 + ba_100, data = covidt2)

summary(covid.carcommute.ba.lm)
```

```{r}
covid.sfh.rent.lm <- lm(covid_case_rate ~ sfh100 + median_gross_rent, data = covidt2)

summary(covid.sfh.rent.lm)
```

```{r}
covid.carcommute.rent.lm <- lm(covid_case_rate ~ carcommute100 + median_gross_rent, data = covidt2)

summary(covid.carcommute.rent.lm)
```

```{r}
covid.carcommute.ba.rent.sfh.lm <- lm(covid_case_rate ~ carcommute100 + ba_100 + median_gross_rent + sfh100, data = covidt2)

summary(covid.carcommute.ba.rent.sfh.lm)
```

```{r}
covid.carcommute.sfh.lm <- lm(covid_case_rate ~ carcommute100 + sfh100, data = covidt2)

summary(covid.carcommute.sfh.lm)
```

```{r}
covid.carcommute.b10.lm <- lm(covid_case_rate ~ carcommute100 + x10_b100, data = covidt2)

summary(covid.carcommute.b10.lm)
```

```{r}
covid.carcommute.ba.rent.b10.lm <- lm(covid_case_rate ~ carcommute100 + ba_100 + median_gross_rent + x10_b100, data = covidt2)

summary(covid.carcommute.ba.rent.b10.lm)
```

```{r}
covid.crowding.b10.lm <- lm(covid_case_rate ~ crowding100 + x10_b100, data = covidt2)

summary(covid.crowding.b10.lm)
```

```{r}
covid.carcommute.ba.rent.lm <- lm(covid_case_rate ~ carcommute100 + ba_100 + median_gross_rent, data = covidt2)

summary(covid.carcommute.ba.rent.lm)
```

```{r}
covid.ba.sfh.lm <- lm(covid_case_rate ~ ba_100 + sfh100, data = covidt2)

summary(covid.ba.sfh.lm)
```

```{r}
covid.crowding.sfh.lm <- lm(covid_case_rate ~ crowding100 + sfh100, data = covidt2)

summary(covid.crowding.sfh.lm)
```

```{r}
covid.crowding.white.lm <- lm(covid_case_rate ~ crowding100 + white100, data = covidt2)

summary(covid.crowding.white.lm)
```

```{r}
covid.ba.rent.lm <- lm(covid_case_rate ~ ba_100 + median_gross_rent, data = covidt2)

summary(covid.ba.rent.lm)
```

Here I discovered that the most important variable was one we didn't have: support for Trump. I tried to do a last-minute proxy for Trump support by just controlling for Staten Island, but that didn't work all that well, obviously.

```{r}
covidt23 <- covidt2 %>%
  mutate(staten_island = case_when(
    startsWith(boro, "Staten Island") ~ 1
  ))
```

```{r}
covidt22 <- covidt2 %>%
  mutate(staten_island = boro == "Staten Island")
```

```{r}
covid.crowding.staten.lm <- lm(covid_case_rate ~ crowding100 + staten_island, data = covidt22)

summary(covid.crowding.staten.lm)
```

```{r}
covid.sfh.staten.lm <- lm(covid_case_rate ~ sfh100 + staten_island, data = covidt22)

summary(covid.sfh.staten.lm)
```

```{r}
covid.sfh.staten.income.lm <- lm(covid_case_rate ~ sfh100 + staten_island + median_household_income, data = covidt22)

summary(covid.sfh.staten.income.lm)
```

```{r}
covid.staten.lm <- lm(covid_case_rate ~ staten_island, data = covidt22)

summary(covid.staten.lm)
```

I modified this ggplot below to get all of the charts that look like they were made in R.

```{r}
ggplot(data = covidt22) +
  aes(x = staten_island, y = covid_case_rate) +
  geom_point(aes(color = boro)) +
  geom_smooth(method='lm', se = FALSE)
```




